Your main manuscript title

### 2021-06-17 forecast/paper/paper.tex
added up/down supplement section

[TODO: should we
insert one line explaining this---you're somewhat protected form volatiliy in a
logistic model thanks to 
the sigmoid function used to define the probabilities?]

@@

While the performance of all forecast degrades during up periods (WIS is on the
scale of cases, so higher case counts means higher WIS) there are pronounced
patterns for the different models. 
Google-AA is the only model that
significantly outperforms the AR model during upswings. Both Google-AA and
CTIS-CLI-in-community have large right tails during the flat periods while the
claims signals have large left tails (indicating some large, relatively poor
forecasts) during flat and up periods. CHNG-CLI, CHNG-COVID, Google-AA and
especially CTIS-CLI-in-community have large positive outliers during the down
period. Overall, the indicators seem to help more during periods of flat or down
cases than during up periods, with the exception of Google-AA. In the
supplement, we examine classification accuracy and log-likelihood for the
hotspot task. We see a similar phenomenon: the indicators dramatically improve
hotspot prediction accuracy during flat and down periods with more mixed
behavior during periods of increasing cases, with CHNG-CLI, CHNG-COVID, and DV-CLI
leading to decreased performance.

@@

>>>>

[TODO: should we
insert one line explaining this---you're somewhat protected form volatiliy in a
logistic model thanks to 
the sigmoid function used to define the probabilities? DJM: I think not, seems a
bit far afield.]

@@

While the performance of all forecasters degrades during up periods (WIS is on the
scale of cases, so higher case counts means higher WIS) there are pronounced
patterns for the different models. 
Google-AA is the only model that
outperforms the AR model, on average, during upswings (see the supplement for
these averaged conclusions). Both Google-AA and
CTIS-CLI-in-community have large right tails during the flat periods. CHNG-CLI
and DV-CLI have large left tails (indicating some large, relatively poor
forecasts) during flat and up periods. CHNG-CLI, CHNG-COVID, Google-AA, and
especially CTIS-CLI-in-community have large positive outliers during the down
period. Overall, the indicators seem to help more during periods of flat or down
cases than during periods of increasing case rates, with the exception of
Google-AA.
In the supplement, we examine classification accuracy and log-likelihood for the
hotspot task. We see a similar phenomenon: the indicators dramatically improve
hotspot prediction accuracy during flat and down periods with more mixed
behavior during periods of increasing cases, with CHNG-CLI, CHNG-COVID, and DV-CLI
leading to decreased performance.

@@

The supplement pursues this analysis further. Essentially, during downswings,
improved performance is strongly correlated with predicing larger declines than
the AR model. The AR model tends to overpredict relative to the decline, but
adding indicators tends to diminish this effect. During upswings, the AR model
tends to underpredict, but when the addition of an indicator improves
performance (a more rare occurrence) it is rarely due to higher median
predictions.

### 2021-06-17 forecast/paper/supplement.tex
added up/down supplement section

The \chngcli~indicator (along with the other claims-based signals) is the most
affected by this distinction, reflecting the latency in claims-based
reporting.  This supports the importance of efforts to provide ``nowcasts'' for claims signals
(which corresponds to a 0-ahead ``forecast'' of what the claims
signal's value will be once all data has been collected). Looking at
the \chngcli~and \dv~curves in Figure \ref{fig:fcast-finalized}, we
can see that they perform very similarly when trained on the finalized data.  This is reassuring
because they are in principle measuring the same thing (namely, the percentage of
  outpatient visits that are primarily about COVID-related symptoms).
  The substantial difference in their curves in Figure 3 of the
  main paper must therefore reflect their having very different
  backfill profiles.

While \dv~is one of the least affected methods by using finalized
rather than vintage values, it is one of the most affected methods for the hotspot problem.
This is a reminder that the forecasting and hotspot problems are
distinct problems.  For example, the hotspot problem does not measure
the ability to distinguish between flat and downward trends.

Even the \ar~model is affected by this distinction, reflecting the
fact that the case rates themselves (i.e., the response values) are also
subject to revision.  The forecasters based on indicators are thus
affected both by revisions to the indicators and by revisions to the
case rates.  In the case of the \gs~model, in which we only used finalized
values for the \gs~indicator, the difference in performance can be
wholly attributed to revisions of case rates.

@@

In this section, we consider using the geometric mean instead of the
arithmetic mean when aggregating the weighted interval score (WIS) across location-time pairs.
There are three reasons why the geometric mean may be desirable.
\begin{enumerate}
\item  WIS is right-skewed, being bounded below by zero and having
  occasional very large values.   Figure\textasciitilde{}\ref{fig:wis-densities} illustrates that the
  densities appear roughly log-Gaussian.  The geometric mean is a
  natural choice in such a context since the relative ordering of forecasters is
  determined by the arithmetic mean of the {\em logarithm} of their WIS
  values.
\item In the main paper, we report the ratio of the mean WIS of a
  forecaster to the mean WIS of the baseline forecaster. Another
  choice could be to take the mean of the ratio of WIS values for the
  two methods. (This latter choice would penalize a method less for
  doing poorly where the baseline forecaster also does poorly.)
Using instead the geometric mean makes the order of aggregation and
scaling immaterial since the ratio of geometric means is the same as
the geometric mean of ratios.
\item If one imagines that a forecaster's WIS is composed of
  multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f}$, then
  taking the ratio of two forecasters' geometric mean WIS values will
  effectively cancel these space-time effects.
\end{enumerate}

Figure\textasciitilde{}\ref{fig:fcast-adjusted} uses the geometric
mean for aggregation.  Comparing this with Figure 3 of the main paper,
we see that the main conclusions are largely unchanged; however,
\chngcli~now does appear better than \ar.  This behavior would be
expected if \chngcli's poor performance is attributable to a
relatively small number of large errors (as opposed to a large number
of moderate errors).  Indeed, Figure 5 of the main paper further
corroborates this, in which we see the heaviest left tails occuring
for \chngcli.

\hypertarget{bootstrap-results}{%
\section{Evaluating the implicit regularization hypothesis}\label{bootstrap-results}}

As explained in Section 2.B. of the main paper, a (somewhat cynical)
hypothesis for why we see benefits in forecasting and hotspot
prediction is that the indicators are not actually providing useful
information but they are instead simply acting as a sort of ``implicit
regularization,''  leading to shrinkage on the autoregressive
coefficients, leading to less volatile predictions.  To investigate
this hypothesis, we consider fitting  ``noise features'' that in truth
should have zero coefficients.  Recall (from the main paper) that at each forecast date, we
train a model on 6,426 location-time pairs.  Indicator models are
based on six features, corresponding to the three autoregressive terms
and the three lagged indicator values.  To form ``noise features''
we resample from other rows.  In particular, at each location $\ell$ and
time $t$, we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
X_{\ell,t-14})$ by the triplet $(X_{\ell^*,t^*}, X_{\ell^*,t^*-7},
X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is a location-time pair
sampled with replacement from the 6,426 location-time pairs.  Figures
\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No
method exhibits a noticeable performance gain over the \ar~method,
leading us to dismiss the implicit regularization hypothesis.

@@

\hypertarget{correlations-with-lagged-actuals}{%
\section{Correlations with lagged
actuals}\label{correlations-with-lagged-actuals}}

Alden's histograms are in Figure\textasciitilde{}\ref{fig:cor-wis-ratio}
and Figure\textasciitilde{}\ref{fig:cor-wis-ratio-m1}.

@@

Logged version of Figure 5 in the manuscript is in
Figure\textasciitilde{}\ref{fig:upswing-histogram-logged}.

See also Table\textasciitilde{}\ref{tab:upswing-corr-table}.

@@

In this section, we investigate the sensitivity of the results to the
period over which we train and evaluate the models.  In the main
paper, we end all evaluation on December 31, 2020.  Figures
\ref{fig:fcast-alldates} -- \ref{fig:hot-alldates} show how the
results would differ if we extended this analysis through \attn{XX,
  XX}, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
the main paper, one sees that as ahead increases most methods now improve
relative to the baseline forecaster. When compared to other methods, \chngcli~appears much better than
it had previously; however, all forecasters other than \chngcov~and
\dv~are performing less well relative to the baseline than before.
These changes are likely due to the differing nature of the pandemic
in 2021, with flat and downward trends much more common than upward
trajectories.  Indeed, the nature of the hotspot prediction problem is
quite different in this period.  With a 21-day training window, it is
common for there to be very few hotspots in training.

@@

@@

\caption{This is one of the correlation plots Alden made. It shows histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in 7dav cases relative to 7 days earlier.}\label{fig:cor-wis-ratio}
\end{figure}

{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-m1-1}

\caption{This is Alden's second set of histograms. Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier}\label{fig:cor-wis-ratio-m1}
\end{figure}

{\centering \includegraphics[width=\textwidth]{fig/upswing-histogram-logged-1}

\caption{Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. }\label{fig:upswing-histogram-logged}
\end{figure}

\begin{table}

\caption{\label{tab:upswing-corr-table}Correlation of the difference in WIS between the AR model with the difference in median predictions. In down periods, improvements in forecast risk are highly correlated with lower median predictions. The opposite is true in up periods. This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model. This conclusion is stronger in down periods then in up periods.}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
udf & CHNG-CLI & CHNG-COVID & CTIS-CLIIC & DV-CLI & Google-AA\\
\midrule
down & 0.79 & 0.80 & 0.83 & 0.81 & 0.81\\
flat & 0.12 & 0.17 & 0.28 & 0.18 & 0.17\\
up & -0.57 & -0.56 & -0.53 & -0.53 & -0.47\\
\bottomrule
\end{tabular}
\end{table}

@@

>>>>

The \chngcli\textasciitilde indicator (along with the other claims-based
signals) is the most affected by this distinction, reflecting the
latency in claims-based reporting. This supports the importance of
efforts to provide ``nowcasts'' for claims signals (which corresponds to
a 0-ahead ``forecast'' of what the claims signal's value will be once
all data has been collected). Looking at the \chngcli\textasciitilde and
\dv\textasciitilde curves in Figure \ref{fig:fcast-finalized}, we can
see that they perform very similarly when trained on the finalized data.
This is reassuring because they are in principle measuring the same
thing (namely, the percentage of outpatient visits that are primarily
about COVID-related symptoms). The substantial difference in their
curves in Figure 3 of the main paper must therefore reflect their having
very different backfill profiles.

While \dv\textasciitilde is one of the least affected methods by using
finalized rather than vintage values, it is one of the most affected
methods for the hotspot problem. This is a reminder that the forecasting
and hotspot problems are distinct problems. For example, the hotspot
problem does not measure the ability to distinguish between flat and
downward trends.

Even the \ar\textasciitilde model is affected by this distinction,
reflecting the fact that the case rates themselves (i.e., the response
values) are also subject to revision. The forecasters based on
indicators are thus affected both by revisions to the indicators and by
revisions to the case rates. In the case of the
\gs\textasciitilde model, in which we only used finalized values for the
\gs\textasciitilde indicator, the difference in performance can be
wholly attributed to revisions of case rates.

@@

In this section, we consider using the geometric mean instead of the
arithmetic mean when aggregating the weighted interval score (WIS)
across location-time pairs. There are three reasons why the geometric
mean may be desirable.

\begin{enumerate}
\item  WIS is right-skewed, being bounded below by zero and having
  occasional very large values.   Figure \ref{fig:wis-densities} illustrates that the
  densities appear roughly log-Gaussian.  The geometric mean is a
  natural choice in such a context since the relative ordering of forecasters is
  determined by the arithmetic mean of the {\em logarithm} of their WIS
  values.
\item In the main paper, we report the ratio of the mean WIS of a
  forecaster to the mean WIS of the baseline forecaster. Another
  choice could be to take the mean of the ratio of WIS values for the
  two methods. (This latter choice would penalize a method less for
  doing poorly where the baseline forecaster also does poorly.)
Using instead the geometric mean makes the order of aggregation and
scaling immaterial since the ratio of geometric means is the same as
the geometric mean of ratios.
\item If one imagines that a forecaster's WIS is composed of
  multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f}$, then
  taking the ratio of two forecasters' geometric mean WIS values will
  effectively cancel these space-time effects.
\end{enumerate}

Figure \ref{fig:fcast-adjusted} uses the geometric mean for aggregation.
Comparing this with Figure 3 of the main paper, we see that the main
conclusions are largely unchanged; however, \chngcli\textasciitilde now
does appear better than \ar. This behavior would be expected if
\chngcli's poor performance is attributable to a relatively small number
of large errors (as opposed to a large number of moderate errors).
Indeed, Figure 5 of the main paper further corroborates this, in which
we see the heaviest left tails occuring for \chngcli.

\hypertarget{bootstrap-results}{%
\section{Bootstrap results}\label{bootstrap-results}}

As explained in Section 2.B. of the main paper, a (somewhat cynical)
hypothesis for why we see benefits in forecasting and hotspot prediction
is that the indicators are not actually providing useful information but
they are instead simply acting as a sort of implicit regularization,"
leading to shrinkage on the autoregressive coefficients, leading to less
volatile predictions. To investigate this hypothesis, we consider
fitting ``noise features'' that in truth should have zero coefficients.
Recall (from the main paper) that at each forecast date, we train a
model on 6,426 location-time pairs. Indicator models are based on six
features, corresponding to the three autoregressive terms and the three
lagged indicator values. To form noise features we resample from other
rows. In particular, at each location \(\ell\) and time \(t\), we
replace the triplet \((X_{\ell,t}, X_{\ell,t-7}, X_{\ell,t-14})\) by the
triplet \((X_{\ell^*,t^*}, X_{\ell^*,t^*-7}, X_{\ell^*,t^*-14})\), where
\((\ell^*,t^*)\) is a location-time pair sampled with replacement from
the 6,426 location-time pairs. Figures
\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results. No method
exhibits a noticeable performance gain over the
\ar\textasciitilde method, leading us to dismiss the implicit
regularization hypothesis.

@@

@@

In this section we provide extra details about the upswing / flat /
downswing analysis described in the main text. Figure
\ref{fig:upswing-summary} shows the overall results, examining the
average difference \(\WIS(AR) - \WIS(F)\) in period. Figure
\ref{fig:hotspots-upswing-downswing} shows the same information for the
hotspot task. On average, during downswings and flat periods, the
indicator-assisted models have lower classification error and higher log
likelihood than the AR model. For hotspots, both Google-AA and
CTIS-CLIIC perform better than the AR model during upswings, in contrast
to the forecasting task, where only Google-AA improves. For a related
analysis, Figure \ref{fig:cor-wis-ratio} shows histograms of the
Spearman correlation (Spearman's \(\rho\), a rank-based measure of
association) between the \(\WIS(F)/\WIS(AR)\) and the magnitude of the
swing. Again we see that case rate increases are positively related to
diminished performance of the indicator models.

One hypothesis for diminished relative performance during upswings is
that the AR model tends to overpredict downswings and underpredict
upswings. Adding indicators seems to help avoid this behavior on the
downswing but not as much on upswings. Figure
\ref{fig:upswing-corr-table} shows the correlation between between
\(\WIS(AR) - \WIS(F)\) and the difference of their median forecasts.
During downswings, this correlation is large, implying that improved
relative performance of \(F\) is related to making lower forecasts than
the AR model. The opposite is true during upswings. This is largely to
be expected. However, the relationship attenuates in flat periods and
during upswings. That is, when performance is better in those cases, it
may be due to other factors than simply making predictions in the
correct direction, for example, narrower confidence intervals.

@@

In this section, we investigate the sensitivity of the results to the
period over which we train and evaluate the models. In the main paper,
we end all evaluation on December 31, 2020. Figures
\ref{fig:fcast-alldates} -- \ref{fig:hot-alldates} show how the results
would differ if we extended this analysis through March 31, 2021.
Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of the main paper,
one sees that as ahead increases most methods now improve relative to
the baseline forecaster. When compared to other methods,
\chngcli\textasciitilde appears much better than it had previously;
however, all forecasters other than \chngcov\textasciitilde and
\dv\textasciitilde are performing less well relative to the baseline
than before. These changes are likely due to the differing nature of the
pandemic in 2021, with flat and downward trends much more common than
upward trajectories. Indeed, the nature of the hotspot prediction
problem is quite different in this period. With a 21-day training
window, it is common for there to be very few hotspots in training.

@@

{\centering \includegraphics[width=\textwidth]{fig/upswing-summary-1}

@@

\caption{Average difference between the WIS of the AR model and the WIS of the other forecasters. The indicator-assisted forecasters do best during down and flat periods.}\label{fig:upswing-summary}
\end{figure}

@@

\caption{Histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in smoothed case rates relative to 7 days earlier.}\label{fig:cor-wis-ratio}
\end{figure}

{\centering \includegraphics[width=\textwidth]{fig/upswing-corr-table-1}

\caption{Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster. In down periods, improvements in forecast risk are highlycorrelated with lower median predictions. The opposite is true in up periods. This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model. This conclusion is stronger in down periods then in up periods.}\label{fig:upswing-corr-table}
\end{figure}
